---
title: "Mash"
author: "Yuxin Zou"
date: 2018-5-10
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(limma); library(mashr); library(mclust); library(plyr);
library(flashr); library(colorRamps); library(corrplot); library(ggplot2)
```

```{r data}
data = readRDS('../data/results/control/CompiledData.rds')
```

## Mash

```{r mash.data}
mash.data = mash_set_data(Bhat = data$beta, pval = data$p)
```

Top genes:
```{r top genes}
# find strong genes
m.1by1 = mash_1by1(mash.data, alpha=0)
strong = get_significant_results(m.1by1)
Z = mash.data$Bhat/mash.data$Shat
Z.strong = Z[strong,]
# center
Z.center = apply(Z.strong, 2, function(x) x - mean(x))
```

### Estimate covariance structures:

Data Driven:

Flash:
```{r flash}
flash.data = flash_set_data(Z.center)
fmodel = flash(flash.data, greedy = TRUE, backfit = TRUE)
```
```{r flash plot}
factors = flash_get_ldf(fmodel)$f
row.names(factors) = colnames(data$beta)
pve.order = order(flash_get_pve(fmodel), decreasing = TRUE)

par(mfrow=c(1,3))
for(i in pve.order){
  barplot(factors[,i], main=paste0('Factor ',i, ' pve= ', round(flash_get_pve(fmodel)[i],3)), las=2, cex.names = 0.7)
}
par(mfrow=c(1,1))
```

`flash` on the loading:
```{r flashLoading}
loading = fmodel$EL[,1:3]
colnames(loading) = paste0('Factor',seq(1,3))
flash.loading = flash_set_data(loading)
flmodel = flash(flash.loading, greedy = TRUE, backfit = TRUE)
```

The flash prefers the rank 0 model. There is no hidden structure in the loading matrix.

Cluster loadings:
```{r cluster loading}
mod = Mclust(loading)
summary(mod$BIC)
```

Using clustering result to fit `mash`:

$$l_{i}\sim \sum_{j=1}^{m}N(\mu_{j}, \Sigma_{j})$$
We estimate the covariance as $F(\Sigma_j + \mu_{j}\mu_{j}')F'$.
```{r cov from flash}
U_list = alply(mod$parameters$variance$sigma,3)
mu_list = alply(mod$parameters$mean,2)
ll = list()
for (i in 1:length(U_list)){
  ll[[i]] = U_list[[i]] + mu_list[[i]] %*% t(mu_list[[i]])
}

Factors = fmodel$EF[,1:3]
U.loading = lapply(ll, function(U){Factors %*% (U %*% t(Factors))})
names(U.loading) = paste0('Load', "_", (1:length(U.loading)))

# rank 1
Flash_res = flash_get_lf(fmodel)
U.Flash = c(mashr::cov_from_factors(t(as.matrix(factors)), "Flash"), 
            list("tFlash" = t(Flash_res) %*% Flash_res / nrow(Z.center)))
```

PCA:
```{r pca}
U.pca = cov_pca(mash_set_data(Z.center), 3)
```

Canonical
```{r Canonical}
U.c = cov_canonical(mash_set_data(Z.center))
```

Extreme Deconvolution
```{r ED}
U.dd = c(U.pca, U.loading, U.Flash, list('XX' = t(Z.center) %*% Z.center / nrow(Z.center)))

U.ed = cov_ed(mash.data, U.dd, strong)
```

### Mash model:

```{r mash}
mash.model = mash(mash.data, c(U.c, U.ed))
```

## Result

The log-likelihood of fit is
```{r loglik} 
get_loglik(mash.model)
```

Here is a plot of weights learned:
```{r estimated weight}
options(repr.plot.width=12, repr.plot.height=4)
barplot(get_estimated_pi(mash.model), las = 2, cex.names = 0.7)
```

Check XX covariance matrix:

```{r XX}
x           <- mash.model$fitted_g$Ulist[["ED_XX"]]
colnames(x) <- colnames(data$beta)
rownames(x) <- colnames(x)
corrplot(x, method='color', cl.lim=c(-0.2,1), type='upper', addCoef.col = "black", tl.col="black", tl.srt=45, col=colorRampPalette(c("blue","white","red"))(200))
```

```{r XX pca}
layout(matrix(c(1,2,3,4), 2, 2, byrow=TRUE))
svd.out = svd(mash.model$fitted_g$Ulist[["ED_XX"]])
v = svd.out$v
colnames(v) = colnames(data$beta)
rownames(v) = colnames(v)
options(repr.plot.width=10, repr.plot.height=5)
for (j in 1:4)
  barplot(v[,j]/v[,j][which.max(abs(v[,j]))], cex.names = 0.7,
          las = 2, main = paste0("EigenVector ", j, " for XX"))
```

There are `r length(get_significant_results(mash.model))` diferentially expressed genes.

Check pairwise sharing by sign:
```{r}
x = get_pairwise_sharing(mash.model, factor=0)
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- colnames(data$beta)
rownames(x) <- colnames(x)
corrplot.mixed(x, tl.pos="d",upper='color', cl.lim=c(0,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               tl.cex=1.2)
```

Check pairwise sharing by magnitude and sign:
```{r}
x = get_pairwise_sharing(mash.model)
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- colnames(data$beta)
rownames(x) <- colnames(x)
corrplot.mixed(x, tl.pos="d",upper='color', cl.lim=c(0,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               tl.cex=1.2)
```

* Effects compare with SCZ

```{r}
CompareSCZ = diag(5)
CompareSCZ[,2] = -1
CompareSCZ = CompareSCZ[-2,]
row.names(CompareSCZ) = colnames(data$beta)[-2]
```

```{r}
mash.model.SCZ = mash.model
mash.model.SCZ$result = mash_compute_posterior_matrices(mash.model, mash.data, A=CompareSCZ, algorithm.version = 'R')
```

Check pairwise sharing by sign:
```{r}
x = get_pairwise_sharing(mash.model.SCZ, factor=0)
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- row.names(CompareSCZ)
rownames(x) <- colnames(x)
corrplot.mixed(x, tl.pos="d",upper='color', cl.lim=c(0,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               tl.cex=1.2)
```

Check pairwise sharing by magnitude and sign:
```{r}
x = get_pairwise_sharing(mash.model.SCZ)
x[x > 1]    <- 1
x[x < -1]   <- -1
colnames(x) <- row.names(CompareSCZ)
rownames(x) <- colnames(x)
corrplot.mixed(x, tl.pos="d",upper='color', cl.lim=c(0,1), upper.col=colorRampPalette(rev(c("#D73027","#FC8D59","#FEE090","#FFFFBF",
                               "#E0F3F8","#91BFDB","#4575B4")))(40),
               tl.cex=1.2)
```


